---
title: "icdConvert - How to Use the Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{icdConvert - How to Use the Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installing `icdConvert`
The package can  be installed from github with the `devtools` package using the following command:
```{r install, eval = FALSE}
devtools::install_github("https://github.com/jradack/icdConvert")
```
You might get a warning about needing Rtools to build the package, but you should be able to ignore this.

Then it can be loaded like any other R package:
```{r setup}
library(icdConvert)
```

The package contains three primary functions(`map_code`, `get_description`, and `map_describe`) and a dataframe (`example_data_icd9_pc`).
The example dataframe has the character column `icd9_pc`, which is a list of ICD-9 procedure codes.
This vignette will demo these primary functions using `example_data_icd9_pc`.
```{r echo = FALSE, results='asis'}
knitr::kable(example_data_icd9_pc[1:10,])
```

# Function demos
## `map_code`
This function takes a vector or ICD codes and converts them to the corresponding ICD version.
The user must specify the destination ICD version (i.e. which ICD version they want to convert the codes to) and the type of code being converted (`"dg"` for diagnosis codes, `"pc"` for procedure codes).
This package handles one-to-many mappings of codes.
When no match can be found, the returned destination code is the string `"No matches found"`.
```{r}
# map_code_test <- map_code(example_data_icd9_pc$icd9_pc, icdVer_dest = 10, code_type = "pc")
```

```{r echo = FALSE, results = 'asis'}
# knitr::kable(map_code_test[1:10,])
```
The mapping is performed using General Equivalence Mapping (GEM) files provided by [CMS](https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes).
This package uses the 2018 version of the GEMs files which were the last year which the GEMs files were updated (see [here](https://www.cms.gov/Medicare/Coding/ICD10/2019-ICD-10-CM)):

> As stated in the FY 2016 IPPS/LTCH PPS final rule (80 FR 49388), the GEMs have been updated on an annual basis as part of the ICD-10 Coordination and Maintenance Committee meetings process and will continue to be updated for approximately 3 years after ICD-10 is implemented.

## `get_description`
This function takes a list of ICD codes and gets the code descriptions using CMS provided tables.
For consistency, the 2018 version of the tables were used for ICD-10.
```{r}
# get_description_test <- get_description(example_data_icd9_pc$icd9_pc, icdVer = 9, code_type = "pc")
```

```{r echo=FALSE, results='asis'}
# knitr::kable(get_description_test[1:10,])
```

## `map_describe`
This function combines the functionality of `map_code` and `get_description` into one.
It maps the codes to the desired ICD version and gets the corresponding descriptions for the mapped codes.
```{r}
# map_describe_test <- map_describe(example_data_icd9_pc$icd9_pc, icdVer_dest = 10, code_type = "pc")
```

```{r echo=FALSE, results='asis'}
# rownames(map_describe_test) <- NULL
# knitr::kable(map_describe_test[1:10,])
```





